---
title: "p8105_hw2_sl5685"
author: "Shumei Liu"
date: "2024-10-02"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

### Problem 1

```{r}
trans_ent = 
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) |> 
  janitor::clean_names() |> 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, vending, entrance_type, ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

trans_ent
```

This dataset contains 19 variables and 1,868 rows. The variables are line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route 7, route8, route9, route10, route11, entry, vending, entrance_type and ada. 

For the data cleaning, first read csv file and clean the names of each columns. Then, select all the variables that we need and use `mutate` convert data from character to a logical variable. The resulting dataset is 1,868 x 19. These data are not tidy.

###### Distinct station

```{r}
trans_ent |> 
  select(station_name, line) |> 
  distinct()
```

There are 465 distinct stations.

###### ADA compliant

```{r}
trans_ent |> 
  filter(ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```

84 stations are ADA compliant.

###### Proportion without vending

```{r}
trans_ent |> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```

###### Serve the A train

```{r}
trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()
```

###### Serve the A train and ADA compliant

```{r}
trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A", ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```

### Problem 2

Mr. Trash Wheel

```{r}
mr_df = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Mr. Trash Wheel") %>%
  janitor::clean_names() |>
  select(-c(x15, x16)) |>
  drop_na() |>
  mutate(sports_balls = as.integer(round(sports_balls))) |>
  mutate(name = "Mr. Trash Wheel") |>
  relocate(name)

mr_df
```

Professor Trash Wheel

```{r}
prof_df = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Professor Trash Wheel") %>%
  janitor::clean_names() |>
  drop_na() |>
  mutate(year = as.character(year)) |>
  mutate(name = "Professor Trash Wheel") |>
  relocate(name)

prof_df
```

Gwynnda Trash Wheel

```{r}
ga_df = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Gwynnda Trash Wheel") %>%
  janitor::clean_names() |>
  drop_na() |>
  mutate(year = as.character(year)) |>
  mutate(name = "Gwynnda Trash Wheel") |>
  relocate(name)

```

Binding rows

```{r}
wheel_tidy = bind_rows(mr_df, prof_df, ga_df)
  
wheel_tidy
```

After cleaning, the dataset includes a total of `r nrow(wheel_tidy)` observations, with each observation representing a specific dumpster filled by one of the trash wheels. The example of key variables include `date`, `month`, `year`, `weight_tons`, and `cigarette_butts`. For available data, the total weight of trash collected by Professor Trash Wheel was `r sum(prof_df$weight_tons)` tons. The total number of cigarette butts collected by Gwynnda in June of 2022 was.







